---
title: "admission"
author: "Dagim"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Admission data set
# The head funciton returns the first 6 rows but we can cuztumize it as we want for eg if we want the first 2 rows we can use -- head(adm, 2) ---
```{r}
adm = read.csv("admission.csv")
adm


```

# The tail function returns the last 6 rows
The tail function return the last 6 rows but we can cuztumize it to tail(adm, 2) if we want the last 2 rows
```{r}
tail(adm, 2)
```
#str(data) Displays the structure of an object, including its type and the first few elements.
```{r}
str(adm)
```

#summary() : provides summary statistics of each column in the dataset like mean median quartiles and more.
```{r}
summary(adm)
```
#dim(): Retruns the dimentions of data (rows and colomns of data)
(
```{r}
dim(adm)
```

#calculating the mean of one col in the data set

```{r}
mean(adm$apt)
```

#calculating the variance and standard deviation of a col

```{r}
vari <- var(adm$apt)
c("variance: ", vari)
stand <- sd(adm$apt)
c("standard deviation: ", stand)
```
### using vector addition if the Grater vector is a multiple of the shorter object then R will copy the rest of the empty part and out put if not it will throw a warning.
```{r}
u = c(9,3,4)
h = c(4,5,8,6,7,1)
print(u > h)
```
### for loops in r
```{r}
x <- rnorm(5)
for(i in 1:5){
  print(x[i])
}
```
### what is the difference between vectorized approach and non vectorized one in R programming to add to random numbers 

```{r}
n = 10
o <- rnorm(n)
u <- rnorm(n)
# the vectorized will take less time because it does have to translate everything to c programming R is basically a wrapper of c program
c <- o * u
print(c)

#unlike other programming laguages this approach will take longer in R
d <-rep(NA, n)

for(i in 1:n){
  d[i] <- o[i] * u[i] 

}

d

```

### importing text data 

```{r}
co <- read.table("cola.txt")
co